Converting LaTeX R Journal Articles into Rmarkdown using texor and rebib

### 2023-07-13 texor.Rmd
revised intro

The R Journal is the primary open-access outlet for the R community. In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations. One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.

There are numerous benefits of HTML format:

1. Articles can include interactive graphics and tables.
2. The format is more accessible to screen readers making the work more accessible to vision impaired researchers.

@@

Describe changes in templates, and web site. Desirability to convert the legacy articles. Decided to convert latex input to Rmd, rather than directly html and why.

- R Journal latex template and formatting
- Quirks of the legacy articles

>>>>

The R Journal is the primary open-access outlet for publications produced by the R community. It was born in 2008, evolving from a newsletter, that ran from 2001, into a more formal article publication to encourage documenting statistical computing research.

The format is constantly evolving. Early articles were typeset using LateX [@latex], from a specific, but changing, template. This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported. With the emergence of dynamic document systems like R Markdown [@rmarkdown] a tight-coupling of code and documentation is possible. Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} [@knitr], making reporting of computing research more reproducible.

In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations. One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf. There are numerous benefits of HTML format:

1. Articles can include interactive graphics and tables.
2. The format is more accessible to screen readers making the work more accessible to vision-impaired researchers.

@@

This latter point is a reason to consider converting all of the legacy articles into HTML.

A key decision for creating conversion software was to decide to directly convert LaTeX to HTML, or PDF to HTML, or LaTeX to R Markdown, and then use the current journal tools to create the HTML. The latter approach was decided to be the most versatile and useful. If an article can be converted from LaTeX to R Markdown, it would help authors make the transition to reproducible publishing, beyond what the R Journal needed. Once an article is in R Markdown format it can be adapted to include the code for dynamic execution.

In addition to article format, changes to the web site structure were important for delivering the publication. Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows for the journal web site to optimally deliver R Markdown articles.

The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal, and to embed them into the journal web site. The packages described here, \CRANpkg{texor} and \CRANpkg{rebib} describes software to convert legacy LaTeX format articles into Rmarkdown, so that they can be rendered in HTML in the new web site.

The paper is organised as follows. Section XXX gives an overview of the conversion process. Section XXX describes examples of pre-processing using regular expressions, and section XXX provides examples of post-processing using lua filters. Section XXX describes tools to do special handling of bibliography files. Section XXX runs through an example conversion process.

<!--
Describe changes in templates, and web site. Desirability to convert the legacy articles. Decided to convert latex input to Rmd, rather than directly html and why.

- R Journal latex template and formatting
- Quirks of the legacy articles
-->
